Statistical methods for relative quantification of post-translational modifications in global proteomics experiments

### 2021-09-09 supplementary/ptm_sm.tex
Added analysis code and experiment papers

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{TMT experiment}
\label{sec:tmtmethod}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=.85\textwidth]{sim_new/sim1_FDR_all_models}\\
\includegraphics[width=.85\textwidth]{sim_new/sim1_FDR}
\caption{(upper plot)All the considered methods in simulation 1 correctly calibrated the Type I error rate when adjusting for changes in protein abundance. In comparison, the methods without accounting for the protein-level changes resulted in off-target, high false positive rates. (lower plot) The considered methods with protein adjustment are compared in detail. The proposed method generally performed better than Limma, but had a higher FDR than T-test. All methods performed strongly, with all FDR rates near or below .03. \label{fig:sim1_fpr}}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=.85\textwidth]{sim_new/sim1_Recall}\\
\includegraphics[width=.85\textwidth]{sim_new/sim1_accuracy}
\caption{(upper plot) The recall for simulation 1 methods with adjustment were compared. It is clear that Limma performs the strongest here when the number of replicates were low.  At higher replicates the performance of the proposed methods and Limma are comparable. T-test clearly performs worse across all methods. (lower plot) The overall accuracy plot mimics the observations in the recall plot. Limma performs stronger than the proposed method at lower replicates, while at higher they are comparable. \label{fig:sim1_recall}}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=.85\textwidth]{sim_new/sim3_Recall}
\includegraphics[width=.85\textwidth]{sim_new/sim3_Accuracy}
\caption{The advantage of using the proposed approach over Limma and $t$-test is apparent when looking at simulation 2, which includes limited observations and the presence of missing values. (upper plot) In the case of recall the proposed method performs stronger than limma and t-test in nearly every model. Even at lower replicates the proposed method still outperformed limma, especially when there were more conditions in the simulation. Again the lowest performance was observed using the $t$-test method. (lower plot) Again in the overall accuracy plots, the results are similar to recall. The propsed method performed the strongest across all methods, even when replicates were low. While Limma shows strong performance in a clean experiment, when real world data problems are introduced it is clear the proposed method is more robust. \label{fig:sim2_recall}}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=.85\textwidth]{sim_new/sim1_FC_boxplot}
\includegraphics[width=.85\textwidth]{sim_new/sim3_FC_boxplot}
\caption{(upper plot) In simulation 1 all considered methods correctly estimated the fold change between conditions, with a median fold change estimation of .75 in all methods. The distributions around the median were also consistent acorss all methods. Predictably the distribution quantiles decrease as the number of replicates increases. (lower plot) In simulation 2 again all methods correctly estimated the fold change with a median log change of .75. The proposed method in this simulation had a visibly tighter distribution around the median. Both Limma and $t$-test showed a wider range around the fold change. In this case the proposed method showed a stronger performed in correctly estimating the fold change for all peptides. \label{fig:fc_boxplot}}
\end{figure}

@@

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{TMT experiment}
\label{sec:tmtsim}

\todo{add the simulation for TMT experiment, batch vs plex?}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{TMT experiment}
\label{sec:tmtmethod}

@@

\begin{figure}[h!]
\centering
 \begin{subfigure}{.9\textwidth}
	\includegraphics[width=.9\textwidth]{sim_new/sim1_FDR_all_models}\\
	\caption{All the considered methods in simulation 1 correctly calibrated the Type I error rate when adjusting for changes in protein abundance. In comparison, the methods without accounting for the protein-level changes resulted in off-target, high false positive rates.}
 \end{subfigure}
 \begin{subfigure}{.9\textwidth}
	\includegraphics[width=.9\textwidth]{sim_new/sim1_FDR}
	\caption{The considered methods with protein adjustment are compared in detail. The proposed method generally performed better than Limma, but had a higher FDR than T-test. All methods performed strongly, with all FDR rates near or below .03.}
 \end{subfigure}
\label{fig:sim1_fpr}
\end{figure}

\begin{figure}[h!]
\centering
 \begin{subfigure}{.9\textwidth}
	\includegraphics[width=.9\textwidth]{sim_new/sim1_Recall}\\
	\caption{The recall for simulation 1 methods with adjustment were compared. It is clear that Limma performs the strongest here when the number of replicates were low.  At higher replicates the performance of the proposed methods and Limma are comparable. T-test clearly performs worse across all methods.}
 \end{subfigure}
 \begin{subfigure}{.9\textwidth}
	\includegraphics[width=.9\textwidth]{sim_new/sim1_accuracy}
	\caption{The overall accuracy plot mimics the observations in the recall plot. Limma performs stronger than the proposed method at lower replicates, while at higher they are comparable.}
 \end{subfigure}
\label{fig:sim1_recall}
\end{figure}

\begin{figure}[h!]
\centering
 \begin{subfigure}{.9\textwidth}
	\includegraphics[width=.9\textwidth]{sim_new/sim3_Recall}
	\caption{The advantage of using the proposed approach is apparent when looking at simulation 2, which includes limited observations and the presence of missing values. In the case of recall the proposed method performs stronger than limma and t-test in nearly every model. Even at lower replicates the proposed method still outperformed limma, especially when there were more conditions in the simulation. Again the lowest performance was observed using the $t$-test method.}
 \end{subfigure}
 \begin{subfigure}{.9\textwidth}
	\includegraphics[width=.9\textwidth]{sim_new/sim3_Accuracy}
	\caption{The overall accuracy plots show similar results to recall. The proposed method performed the strongest across all methods, even when replicates were low. Limma shows strong performance in a clean experiment, when real world data problems are introduced it is clear the proposed method is more robust.}
 \end{subfigure}
\label{fig:sim2_recall}
\end{figure}

\begin{figure}[h!]
\centering
 \begin{subfigure}{.9\textwidth}
	\includegraphics[width=.9\textwidth]{sim_new/sim1_FC_boxplot}
	\caption{In simulation 1 all considered methods correctly estimated the fold change between conditions, with a median fold change estimation of .75 in all methods. The distributions around the median were also consistent acorss all methods. Predictably the distribution quantiles decrease as the number of replicates increases.}
 \end{subfigure}
 \begin{subfigure}{.9\textwidth}
	\includegraphics[width=.85\textwidth]{sim_new/sim3_FC_boxplot}
	\caption{In simulation 2 again all methods correctly estimated the fold change with a median log change of .75. The proposed method in this simulation had a visibly tighter distribution around the median. Both Limma and $t$-test showed a wider range around the fold change. In this case the proposed method showed a stronger performed in correctly estimating the fold change for all peptides.}
 \end{subfigure}
\label{fig:fc_boxplot}
\end{figure}

@@

@@

\clearpage

\subsection{Shigella flexneri}

In this study, the correlation between the autophagy gene Atg16L1 and killing of virulent Shigella flexneri (S.flexneri) was assessed. It was discovered that the loss of Atg16L1 in macrophages enhanced the killing of S.flexneri. Multiplex proteomics was used to quantify the abundance of total protein, phosphorylation, and ubiquitination in wild type (WT) and ATG16L1-deficient (cKO) bone marrow macrophages, uninfected and uninfected with S.flexneri. The abundance of total protein and post-translation modifications were quantified at three time points, uninfected, early infections, and late infection. Quantifying the total protein along with the post-translational modifications allows us to adjust for changes in total protein and see the true impact of the site specific modifications.
